---
title: "Color Guide"
author: "Wanding Zhou"
date: "May 25, 2016"
output: html_document
---

This document illustrate how to use color in WHeatmap.

```{r setup, include=FALSE}
## library(devtools)
## load_all('~/tools/wheatmap/wheatmap')
```

```{r prepare}
library(wheatmap)
```

# continous color map

Usually wheatmap sets colors automatically
```{r}
WColorBarH(1:10)
```

# discrete color map

Even if you specify the color name, it just ignore them and assign its own colors. 
```{r}
a <- WColorBarV(rep(c('black','red','blue'),3))
a
```

One can override by telling WHeatmap the data is color name using the `cmp=CMPar(use.data=T)` option.
```{r}
a + WColorBarV(rep(c('black','red','blue'),3),name='a',RightOf(),cmp=CMPar(use.data=T)) + WLegendV(NULL, BottomRightOf(h.pad=0.05), height=0.1)
```

One can also specify a color mapping manually using the option `cmp=CMPar(label2color=`:
```{r}
a + WColorBarV(rep(c('Fred','Frank','Froy'),3),RightOf(),cmp=CMPar(label2color=c(Fred='green',Frank='red',Froy='yellow'))) + WLegendV(NULL, BottomRightOf(h.pad=0.05), height=0.1)
```