---
title: "tutorial of wheatmap"
output: pdf_document
---

This is a tutorial on the usage of wheatmap for generating complex heatmaps in a procedure way.

A heatmap is easy to generate in R. There are many packages like heatmap2, heatmap3, heatmap.plus and ComplexHeatmaps. But many times, I found myself in a situation of plotting a set of heatmaps in a procedure way. And none-of the above packages are nimble enough to do color-bar reordering, multiple heatmaps that can be positioned arbitrarily. So I wrote this package for that purpose.

We start with some data
```{r}
library(devtools)
load_all('~/tools/wheatmap/wheatmap/')
m <- cbind(matrix(rnorm(20),nrow=4), 5+matrix(rnorm(8),nrow=4))
m2 <- matrix(1:16,nrow=4)
dimnames(m) <- list(c('w','x','y','z'), c('a','b','c','d','e','f','g'))
row.data <- c(1,2,3,1)
col.data <- c(1:6,6)
m
```

We perform some clustering
```{r}
c <- both.cluster(m)
row.data <- row.data[c$row.clust$order]
col.data <- col.data[c$column.clust$order]
```

We plot
```{r}
load_all('~/tools/wheatmap/wheatmap/')

a <- WHeatmap(c$mat, name='h1')

a <- a + WColorBarV(row.data, LeftOf('h1'), 'c1')

a <- a + WColorBarH(col.data, TopOf('h1'), 'c2')

a <- a + WDendrogram(c$row.clust, LeftOf('c1'), facing='right')

a <- a + WDendrogram(c$column.clust, TopOf('c2'), facing='bottom')
  
a <- a + WColorBarV(1:4, RightOf('h1'), 'c3', continuous=TRUE)

a <- a + WHeatmap(m2, RightOf('c3'), 'h2')

a <- a + WColorBarH(rep(c(1,2,3),each=4), Beneath(WGroupColumn('h1', 'c3', 'h2')), 'c4', cmp=CMPar(brewer.name='Set2'), continuous=FALSE)

a <- a + WHeatmap(matrix(rep(c(8:1,1:8),4),nrow=4), Beneath('c4', h.aln=WGroupColumn('h1','c3')), 'h3')

a <- a + WHeatmap(matrix(1:4,nrow=2), RightOf('h3', h.scale='h2'), 'h4')

a <- a + WHeatmap(matrix(1:24,nrow=3), Beneath('h3'), 'h5')

a <- a + WHeatmap(matrix(24:1,nrow=2), Beneath('h5', h.aln=WGroupColumn('h1','c3','h2')), 'h6')

a <- a + WLegendV('c1', RightOf('h6.3'), 'l1')

a <- a + WLegendV('c2', TopOf('l1', pad=0.1), 'l2')

a <- a + WLegendV('c3', TopOf('l2', pad=0.1), n.text=3)

print(a, layout.only)

WPlot(p)
```

Another example of too many labels
```{r}
load_all('~/tools/wheatmap/wheatmap/')
m <- matrix(1:1000, nrow=100)
rownames(m) <- paste0('row', 1:100)
WHeatmap(m, yticklabels = TRUE)
```

